---
title: "p8105_hw2_rh2916"
author: "Rui Huang"
date: "September 29, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(skimr)
library(dplyr)
library(readxl)
```

## Problem 1 

## Read and manipulate Data 

```{r}
Problem1_data =
  read_csv(file = './data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1: route11, 
         entry, vending, entrance_type, ada) %>%
  mutate(
    entry = recode(entry, "YES" = TRUE, "NO" = FALSE)
  )
```

The Problem1_data dataset contains 19 variables, they are line, station_name, station_latitude, station_ongitude, routes1:route11, entry, vending, entrance type and ada. After reading the dataset, I first clean the variable names using clean_names function, then retain the variables I need by select function, and convert the entry variable from character to a logical variable by mutate and recode function. The dimension of the resulting dataset is `r dim(Problem1_data)` (rows x columns). It is not tidy since the route name and route number are messy.

